---
title: "Pathway enrichment analysis, GEO52158"
author: "German Novakovskiy"
date: "August 21, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
library(knitr)
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(GEOquery)))
suppressWarnings(suppressMessages(library(data.table)))
suppressWarnings(suppressMessages(library(reshape2)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(cowplot)))
suppressWarnings(suppressMessages(library(limma)))
suppressWarnings(suppressMessages(library(tibble)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(edgeR)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(VennDiagram)))
suppressMessages(suppressWarnings(library(ermineR)))
suppressMessages(suppressWarnings(library(hgu133a.db)))
suppressMessages(suppressWarnings(library(ReactomePA)))
suppressMessages(suppressWarnings(library(clusterProfiler)))
suppressMessages(suppressWarnings(library(fgsea)))
suppressMessages(suppressWarnings(library(hgu133plus2.db)))
```

This is an analysis of microarray data, which mostly will not be considered.


## Analysis of first day of differentiation 0h VS 24 h (formation of anterior Primitive streak, mesodendoderm)

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE52158/DEgenes_0h_24h_52158.Rdata")
```

Converting probe ids to gene names:
```{r}
x <- hgu133plus2SYMBOL
# Get the probe identifiers - gene symbol mappings
mapped_probes <- mappedkeys(x)
# Convert to a dataframe
genesym.probeid <- as.data.frame(x[mapped_probes])
head(genesym.probeid)
```

Sorted log Fold Changes give us a ranked list:
```{r}
probes_to_genes <- genesym.probeid %>%
  filter(probe_id %in% rownames(DEgenes_0h_24h_52158))

topProbes <- DEgenes_0h_24h_52158 %>%
  rownames_to_column("probes") %>%
  filter(probes %in% probes_to_genes$probe_id)

probes_to_genes <- probes_to_genes %>% column_to_rownames('probe_id')
symbs <- probes_to_genes[topProbes$probes,]
topProbes$Symbol <- symbs

#for ermineR
ermineInputProbeScores <- topProbes %>% 
  #as.data.frame() %>%
  mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeScores$Symbol))
ermineInputProbeScores <- ermineInputProbeScores[-repeats,]

rownames(ermineInputProbeScores) <- NULL
ermineInputProbeScores <- ermineInputProbeScores %>% 
  column_to_rownames("Symbol")

#for fgsea
ermineInputProbeFGSEA <- topProbes %>% 
  #as.data.frame() %>%
  #mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeFGSEA$Symbol))
ermineInputProbeFGSEA <- ermineInputProbeFGSEA[-repeats,]

rownames(ermineInputProbeFGSEA) <- NULL
ermineInputProbeFGSEA <- ermineInputProbeFGSEA %>% 
  column_to_rownames("Symbol")

scoresFGSEA <- ermineInputProbeFGSEA$logFC
names(scoresFGSEA) <- rownames(ermineInputProbeFGSEA)
```

### Reactome
#### ErmineR Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineReactome <- enrichmentResultReactome$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineReactome)` enriched Reactome pathways with ermineR

#### FGSEA Reactome pathways

```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysReactome)` up-regulated and `r nrow(downPathwaysReactome)` down-regulated Reactome pathways with fgsea

```{r}
upPathwaysReactome %>% kable()
```


### KEGG pathways
#### ErmineR KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineKEGG <- enrichmentResultKEGG$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineKEGG)` enriched KEGG pathways with ermineR. 

#### FGSEA KEGG pathways

```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysKEGG)` up-regulated and `r nrow(downPathwaysKEGG)` down-regulated Reactome pathways with fgsea.

```{r}
upPathwaysKEGG %>% kable()
```

## Analysis of second-fourth days of differentiation 24h VS 96h (formation of definitive endoderm from APS)

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE52158/DEgenes_24h_96h_52158.Rdata")
```

Sorted log Fold Changes give us a ranked list:
```{r}
probes_to_genes <- genesym.probeid %>%
  filter(probe_id %in% rownames(DEgenes_24h_96h_52158))

topProbes <- DEgenes_24h_96h_52158 %>%
  rownames_to_column("probes") %>%
  filter(probes %in% probes_to_genes$probe_id)

probes_to_genes <- probes_to_genes %>% column_to_rownames('probe_id')
symbs <- probes_to_genes[topProbes$probes,]
topProbes$Symbol <- symbs

#for ermineR
ermineInputProbeScores <- topProbes %>% 
  #as.data.frame() %>%
  mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeScores$Symbol))
ermineInputProbeScores <- ermineInputProbeScores[-repeats,]

rownames(ermineInputProbeScores) <- NULL
ermineInputProbeScores <- ermineInputProbeScores %>% 
  column_to_rownames("Symbol")

#for fgsea
ermineInputProbeFGSEA <- topProbes %>% 
  #as.data.frame() %>%
  #mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeFGSEA$Symbol))
ermineInputProbeFGSEA <- ermineInputProbeFGSEA[-repeats,]

rownames(ermineInputProbeFGSEA) <- NULL
ermineInputProbeFGSEA <- ermineInputProbeFGSEA %>% 
  column_to_rownames("Symbol")

scoresFGSEA <- ermineInputProbeFGSEA$logFC
names(scoresFGSEA) <- rownames(ermineInputProbeFGSEA)

```

### Reactome
#### ErmineR Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineReactome <- enrichmentResultReactome$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineReactome)` enriched Reactome pathways with ermineR

#### FGSEA Reactome pathways

```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysReactome)` up-regulated and `r nrow(downPathwaysReactome)` down-regulated Reactome pathways with fgsea

```{r}
upPathwaysReactome %>% kable()
```


### KEGG pathways
#### ErmineR KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineKEGG <- enrichmentResultKEGG$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineKEGG)` enriched KEGG pathways with ermineR. 

#### FGSEA KEGG pathways

```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysKEGG)` up-regulated and `r nrow(downPathwaysKEGG)` down-regulated Reactome pathways with fgsea.

```{r}
upPathwaysKEGG %>% kable()
```

## Analysis of whole differentiation process 0h VS 96 h (formation of Definitive endoderm from hESC)

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE52158/DEgenes_0h_96h_52158.Rdata")
```

Sorted log Fold Changes give us a ranked list:
```{r}
probes_to_genes <- genesym.probeid %>%
  filter(probe_id %in% rownames(DEgenes_0h_96h_52158))

topProbes <- DEgenes_0h_96h_52158 %>%
  rownames_to_column("probes") %>%
  filter(probes %in% probes_to_genes$probe_id)

probes_to_genes <- probes_to_genes %>% column_to_rownames('probe_id')
symbs <- probes_to_genes[topProbes$probes,]
topProbes$Symbol <- symbs

#for ermineR
ermineInputProbeScores <- topProbes %>% 
  #as.data.frame() %>%
  mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeScores$Symbol))
ermineInputProbeScores <- ermineInputProbeScores[-repeats,]

rownames(ermineInputProbeScores) <- NULL
ermineInputProbeScores <- ermineInputProbeScores %>% 
  column_to_rownames("Symbol")

#for fgsea
ermineInputProbeFGSEA <- topProbes %>% 
  #as.data.frame() %>%
  #mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeFGSEA$Symbol))
ermineInputProbeFGSEA <- ermineInputProbeFGSEA[-repeats,]

rownames(ermineInputProbeFGSEA) <- NULL
ermineInputProbeFGSEA <- ermineInputProbeFGSEA %>% 
  column_to_rownames("Symbol")

scoresFGSEA <- ermineInputProbeFGSEA$logFC
names(scoresFGSEA) <- rownames(ermineInputProbeFGSEA)

```

### Reactome
#### ErmineR Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineReactome <- enrichmentResultReactome$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineReactome)` enriched Reactome pathways with ermineR

#### FGSEA Reactome pathways

```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysReactome)` up-regulated and `r nrow(downPathwaysReactome)` down-regulated Reactome pathways with fgsea

```{r}
upPathwaysReactome %>% kable()
```


### KEGG pathways
#### ErmineR KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineKEGG <- enrichmentResultKEGG$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineKEGG)` enriched KEGG pathways with ermineR. 

#### FGSEA KEGG pathways

```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysKEGG)` up-regulated and `r nrow(downPathwaysKEGG)` down-regulated Reactome pathways with fgsea.

```{r}
upPathwaysKEGG %>% kable()
```
