---
title: "Pathway enrichment analysis, GEO52158"
author: "German Novakovskiy"
date: "August 21, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
library(knitr)
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(GEOquery)))
suppressWarnings(suppressMessages(library(data.table)))
suppressWarnings(suppressMessages(library(reshape2)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(cowplot)))
suppressWarnings(suppressMessages(library(limma)))
suppressWarnings(suppressMessages(library(tibble)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(edgeR)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(VennDiagram)))
suppressMessages(suppressWarnings(library(ermineR)))
suppressMessages(suppressWarnings(library(hgu133a.db)))
suppressMessages(suppressWarnings(library(ReactomePA)))
suppressMessages(suppressWarnings(library(clusterProfiler)))
suppressMessages(suppressWarnings(library(fgsea)))
suppressMessages(suppressWarnings(library(hgu133plus2.db)))
```

## Pathways

```{r}
x <- scan("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv", what = "", sep = "\n")
x <- strsplit(x, "[ \t]+")
max.col <- max(sapply(x, length))

## specify col.names as ?read.table suggests
cn <- paste("V", 1:max.col, sep = "")
reactome_pathways <- read.table("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv", fill = TRUE, 
                 col.names = cn, sep = '\t', quote = "")

#reactome_pathways[1:5, 1:10] %>% kable()
```

## Analysis of GSE52158

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE52158/DEgenes_0h_96h_52158.Rdata")
```

Converting probe ids to gene names:
```{r}
x <- hgu133plus2SYMBOL
# Get the probe identifiers - gene symbol mappings
mapped_probes <- mappedkeys(x)
# Convert to a dataframe
genesym.probeid <- as.data.frame(x[mapped_probes])
head(genesym.probeid)
```
```{r}
probes_to_genes <- genesym.probeid %>%
  filter(probe_id %in% rownames(DEgenes_0h_96h_52158))

topProbes <- DEgenes_0h_96h_52158 %>%
  rownames_to_column("probes") %>%
  filter(probes %in% probes_to_genes$probe_id)

probes_to_genes <- probes_to_genes %>% column_to_rownames('probe_id')
symbs <- probes_to_genes[topProbes$probes,]
topProbes$Symbol <- symbs

ermineInputProbeScores <- topProbes %>% 
  #as.data.frame() %>%
  mutate(absolute_logFC = logFC) %>% 
  dplyr::select(Symbol, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) 

#randomly delete repeats (we will delete the lowerst ones)
repeats <- which(duplicated(ermineInputProbeScores$Symbol))
ermineInputProbeScores <- ermineInputProbeScores[-repeats,]

rownames(ermineInputProbeScores) <- NULL
ermineInputProbeScores <- ermineInputProbeScores %>% 
  column_to_rownames("Symbol")

head(ermineInputProbeScores, 10)
```

#### Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv",
                               minClassSize = 15,
                               maxClassSize = 300)

enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
#### KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputProbeScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```

#### Wikipathways

```{r}
#converting symbols to EntrezID
geneSymbols <- rownames(ermineInputProbeScores)
geneEntrez <- bitr(geneSymbols, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

ermineInputGeneScoresRPA <- ermineInputProbeScores %>%
  rownames_to_column("SYMBOL") %>%
  filter(SYMBOL %in% geneEntrez$SYMBOL)

inputGSEA <- merge(ermineInputGeneScoresRPA, geneEntrez, sort=FALSE)

inputGSEA <- inputGSEA %>%
  dplyr::select(absolute_logFC, ENTREZID) %>%
  column_to_rownames('ENTREZID')
```

```{r}
#entrezScores <- as.data.frame(test2)
enrichmentResultWiki <- precRecall(scores = inputGSEA,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

enrichmentResultWiki$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultWiki$results %>% 
  dplyr::select(ID, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("ID", "Corrected p-value", 
                                   "Corrected MF p-value"))
```

#### Trying to run everything with fgsea

```{r}
# scores forfgsea (for absolute)
#ermineInputGeneScoresFGSEA <- DEgenes_0h_96h_75748 %>% 
#  rownames_to_column("gene") %>%
#  #mutate(absolute_logFC = abs(logFC)) %>% 
#  dplyr::select(gene, logFC) %>% 
#  na.omit() %>% 
#  as.data.frame() %>% 
#  arrange(desc(logFC)) %>% 
#  column_to_rownames("gene")


scoresFGSEA <- ermineInputProbeScores$absolute_logFC
names(scoresFGSEA) <- rownames(ermineInputProbeScores)

#for wikipathways
#ermineInputGeneScoresWiki <- ermineInputGeneScoresFGSEA %>%
#  rownames_to_column("SYMBOL") %>%
#  filter(SYMBOL %in% geneEntrez$SYMBOL)

#scoresWikiFGSEA <- merge(ermineInputGeneScoresWiki, geneEntrez, sort=FALSE)

#scoresWikiFGSEA <- scoresWikiFGSEA %>%
#  dplyr::select(logFC, ENTREZID) %>%
#  column_to_rownames('ENTREZID')

scoresWiki <- inputGSEA$absolute_logFC
names(scoresWiki) <- rownames(inputGSEA)
```

Reactome:
```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)
reactomeIdPathway <- reactome_pathways[,c(1,2)]
colnames(reactomeIdPathway) <- c("pathway", "Description") 
fgseaRes <- merge(fgseaRes, reactomeIdPathway, sort = FALSE)

#activated pathways
activPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "Description", "padj", "ES", "NES", "nMoreExtreme"))

#inactivated pathways
inactivPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "Description", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(activPathwaysReactome)` activated and `r nrow(inactivPathwaysReactome)` inactivated pathways in Reactome for this data set.

All active pathways:
```{r}
activPathwaysReactome %>% kable()
```

KEGG:
```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#activated pathways
activPathwaysKEGG <- fgseaRes %>% arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05)

#inactivated pathways
inactivPathwaysKEGG <- fgseaRes %>% arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05)
```

There are `r nrow(activPathwaysKEGG)` activated and `r nrow(inactivPathwaysKEGG)` inactivated pathways in Reactome for this data set.

All active pathways:
```{r}
activPathwaysKEGG
```

WikiPathways:
```{r}
pathwaysWiki <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt")

fgseaRes <- fgsea(pathwaysWiki, scoresWiki, minSize=15, maxSize=300, nperm=10000)

#activated pathways
activPathwaysWiki <- fgseaRes %>% arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05)

#inactivated pathways
inactivPathwaysWiki <- fgseaRes %>% arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05)
```

There are `r nrow(activPathwaysWiki)` activated and `r nrow(inactivPathwaysWiki)` inactivated pathways in Reactome for this data set.

All active pathways:
```{r}
activPathwaysWiki
```

