---
title: "Pathway enrichment analysis, GEO109658"
author: "German Novakovskiy"
date: "August 17, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(knitr)
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(GEOquery)))
suppressWarnings(suppressMessages(library(data.table)))
suppressWarnings(suppressMessages(library(reshape2)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(cowplot)))
suppressWarnings(suppressMessages(library(limma)))
suppressWarnings(suppressMessages(library(tibble)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(edgeR)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(VennDiagram)))
suppressMessages(suppressWarnings(library(ermineR)))
suppressMessages(suppressWarnings(library(hgu133a.db)))
suppressMessages(suppressWarnings(library(ReactomePA)))
suppressMessages(suppressWarnings(library(clusterProfiler)))
suppressMessages(suppressWarnings(library(fgsea)))
```


## Analysis of first day of differentiation 0h VS 24 h (formation of anterior Primitive streak, mesodendoderm)

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE109658/DEgenes_0h_24h_109658.Rdata")
```

Sorted log Fold Changes give us a ranked list:
```{r}
#absolute Log fold changes for ermineR
ermineInputGeneScores <- DEgenes_0h_24h_109658 %>% 
  rownames_to_column("gene") %>%
  mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) %>% 
  column_to_rownames("gene")

#exact log fold changes for fgsea
# scores forfgsea
scoresFGSEADF <- DEgenes_0h_24h_109658 %>% 
  rownames_to_column("gene") %>%
  #mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) %>% 
  column_to_rownames("gene")


scoresFGSEA <- scoresFGSEADF$logFC
names(scoresFGSEA) <- rownames(scoresFGSEADF)

#for WikiPathways
eg <-  bitr(rownames(ermineInputGeneScores), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
ermineInputGeneScoresWiki <- ermineInputGeneScores %>% rownames_to_column("SYMBOL") %>% 
  filter(SYMBOL %in% eg$SYMBOL)
ermineInputGeneScoresWiki <- merge(ermineInputGeneScoresWiki, eg, sort=FALSE)
rownames(ermineInputGeneScoresWiki) <- ermineInputGeneScoresWiki$ENTREZID
ermineInputGeneScoresWiki <- ermineInputGeneScoresWiki %>% dplyr::select("absolute_logFC")

scoresFGSEADFWiki <- scoresFGSEADF %>% rownames_to_column("SYMBOL") %>% 
  filter(SYMBOL %in% eg$SYMBOL)
scoresFGSEADFWiki <- merge(scoresFGSEADFWiki, eg, sort=FALSE)
rownames(scoresFGSEADFWiki) <- scoresFGSEADFWiki$ENTREZID
scoresFGSEADFWiki <- scoresFGSEADFWiki %>% dplyr::select("logFC")

scoresFGSEAWiki <- scoresFGSEADFWiki$logFC
names(scoresFGSEAWiki) <- rownames(scoresFGSEADFWiki)
```

### Reactome
#### ErmineR Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineReactome <- enrichmentResultReactome$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineReactome)` enriched Reactome pathways with ermineR

#### FGSEA Reactome pathways

```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysReactome)` up-regulated and `r nrow(downPathwaysReactome)` down-regulated Reactome pathways with fgsea

```{r}
upPathwaysReactome %>% kable()
```
```{r}
upPathwaysReactome109658_024 <- upPathwaysReactome
save(upPathwaysReactome109658_024, file="upPathwaysReactome109658_024.Rdata")
```


### KEGG pathways
#### ErmineR KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineKEGG <- enrichmentResultKEGG$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineKEGG)` enriched KEGG pathways with ermineR. 

#### FGSEA KEGG pathways

```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysKEGG)` up-regulated and `r nrow(downPathwaysKEGG)` down-regulated Reactome pathways with fgsea.

```{r}
upPathwaysKEGG %>% kable()
```
```{r}
upPathwaysKEGG109658_024 <- upPathwaysKEGG
save(upPathwaysKEGG109658_024, file="upPathwaysKEGG109658_024.Rdata")
```

### WikiPathways
#### ErmineR Wiki pathways

```{r}
enrichmentResultWiki <- precRecall(scores = ermineInputGeneScoresWiki,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways_changed_column.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultWiki$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  filter(CorrectedMFPvalue <= 0.05) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineWiki <- enrichmentResultWiki$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineWiki)` enriched Wiki pathways with ermineR

#### FGSEA Wiki pathways

```{r}
pathwaysWiki <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysWiki, scoresFGSEAWiki, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysWiki <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysWiki <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysWiki)` up-regulated and `r nrow(downPathwaysWiki)` down-regulated Wiki pathways with fgsea

```{r}
upPathwaysWiki %>% kable()
```
```{r}
upPathwaysWiki109658_024 <- upPathwaysWiki
save(upPathwaysWiki109658_024, file="upPathwaysWiki109658_024.Rdata")
```

## Analysis of second-fourth days of differentiation 24h VS 96h (formation of definitive endoderm from APS)

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE109658/DEgenes_24h_96h_109658.Rdata")
```

Sorted log Fold Changes give us a ranked list:
```{r}
#absolute Log fold changes for ermineR
ermineInputGeneScores <- DEgenes_24h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) %>% 
  column_to_rownames("gene")

#exact log fold changes for fgsea
# scores forfgsea
scoresFGSEADF <- DEgenes_24h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  #mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) %>% 
  column_to_rownames("gene")


scoresFGSEA <- scoresFGSEADF$logFC
names(scoresFGSEA) <- rownames(scoresFGSEADF)

#for WikiPathways
eg <-  bitr(rownames(ermineInputGeneScores), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
ermineInputGeneScoresWiki <- ermineInputGeneScores %>% rownames_to_column("SYMBOL") %>% 
  filter(SYMBOL %in% eg$SYMBOL)
ermineInputGeneScoresWiki <- merge(ermineInputGeneScoresWiki, eg, sort=FALSE)
rownames(ermineInputGeneScoresWiki) <- ermineInputGeneScoresWiki$ENTREZID
ermineInputGeneScoresWiki <- ermineInputGeneScoresWiki %>% dplyr::select("absolute_logFC")

scoresFGSEADFWiki <- scoresFGSEADF %>% rownames_to_column("SYMBOL") %>% 
  filter(SYMBOL %in% eg$SYMBOL)
scoresFGSEADFWiki <- merge(scoresFGSEADFWiki, eg, sort=FALSE)
rownames(scoresFGSEADFWiki) <- scoresFGSEADFWiki$ENTREZID
scoresFGSEADFWiki <- scoresFGSEADFWiki %>% dplyr::select("logFC")

scoresFGSEAWiki <- scoresFGSEADFWiki$logFC
names(scoresFGSEAWiki) <- rownames(scoresFGSEADFWiki)
```

### Reactome
#### ErmineR Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineReactome <- enrichmentResultReactome$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineReactome)` enriched Reactome pathways with ermineR

#### FGSEA Reactome pathways

```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysReactome)` up-regulated and `r nrow(downPathwaysReactome)` down-regulated Reactome pathways with fgsea

```{r}
upPathwaysReactome %>% kable()
```
```{r}
upPathwaysReactome109658_2496 <- upPathwaysReactome
save(upPathwaysReactome109658_2496, file="upPathwaysReactome109658_2496.Rdata")
```

### KEGG pathways
#### ErmineR KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineKEGG <- enrichmentResultKEGG$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineKEGG)` enriched KEGG pathways with ermineR. 

#### FGSEA KEGG pathways

```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysKEGG)` up-regulated and `r nrow(downPathwaysKEGG)` down-regulated Reactome pathways with fgsea.

```{r}
upPathwaysKEGG %>% kable()
```
```{r}
upPathwaysKEGG109658_2496 <- upPathwaysKEGG
save(upPathwaysKEGG109658_2496, file="upPathwaysKEGG109658_2496.Rdata")
```

### WikiPathways
#### ErmineR Wiki pathways

```{r}
enrichmentResultWiki <- precRecall(scores = ermineInputGeneScoresWiki,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways_changed_column.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultWiki$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  filter(CorrectedMFPvalue <= 0.05) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineWiki <- enrichmentResultWiki$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineWiki)` enriched Wiki pathways with ermineR

#### FGSEA Wiki pathways

```{r}
pathwaysWiki <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysWiki, scoresFGSEAWiki, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysWiki <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysWiki <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysWiki)` up-regulated and `r nrow(downPathwaysWiki)` down-regulated Wiki pathways with fgsea

```{r}
upPathwaysWiki %>% kable()
```
```{r}
upPathwaysWiki109658_2496 <- upPathwaysWiki
save(upPathwaysWiki109658_2496, file="upPathwaysWiki109658_2496.Rdata")
```

## Analysis of whole differentiation process 0h VS 96 h (formation of Definitive endoderm from hESC)

Load DE data:
```{r}
load("~/ESC_RNA_seq/analysis_of_public_data/GSE109658/DEgenes_0h_96h_109658.Rdata")
```

Sorted log Fold Changes give us a ranked list:
```{r}
#absolute Log fold changes for ermineR
ermineInputGeneScores <- DEgenes_0h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) %>% 
  column_to_rownames("gene")

#exact log fold changes for fgsea
# scores forfgsea
scoresFGSEADF <- DEgenes_0h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  #mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) %>% 
  column_to_rownames("gene")


scoresFGSEA <- scoresFGSEADF$logFC
names(scoresFGSEA) <- rownames(scoresFGSEADF)

#for WikiPathways
eg <-  bitr(rownames(ermineInputGeneScores), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
ermineInputGeneScoresWiki <- ermineInputGeneScores %>% rownames_to_column("SYMBOL") %>% 
  filter(SYMBOL %in% eg$SYMBOL)
ermineInputGeneScoresWiki <- merge(ermineInputGeneScoresWiki, eg, sort=FALSE)
rownames(ermineInputGeneScoresWiki) <- ermineInputGeneScoresWiki$ENTREZID
ermineInputGeneScoresWiki <- ermineInputGeneScoresWiki %>% dplyr::select("absolute_logFC")

scoresFGSEADFWiki <- scoresFGSEADF %>% rownames_to_column("SYMBOL") %>% 
  filter(SYMBOL %in% eg$SYMBOL)
scoresFGSEADFWiki <- merge(scoresFGSEADFWiki, eg, sort=FALSE)
rownames(scoresFGSEADFWiki) <- scoresFGSEADFWiki$ENTREZID
scoresFGSEADFWiki <- scoresFGSEADFWiki %>% dplyr::select("logFC")

scoresFGSEAWiki <- scoresFGSEADFWiki$logFC
names(scoresFGSEAWiki) <- rownames(scoresFGSEADFWiki)

```

### Reactome
#### ErmineR Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineReactome <- enrichmentResultReactome$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineReactome)` enriched Reactome pathways with ermineR

#### FGSEA Reactome pathways

```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Human_Reactome_August_01_2018_symbol.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysReactome <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysReactome)` up-regulated and `r nrow(downPathwaysReactome)` down-regulated Reactome pathways with fgsea

```{r}
upPathwaysReactome %>% kable()
```
```{r}
upPathwaysReactome109658_096 <- upPathwaysReactome
save(upPathwaysReactome109658_096, file="upPathwaysReactome109658_096.Rdata")
```

### KEGG pathways
#### ErmineR KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineKEGG <- enrichmentResultKEGG$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineKEGG)` enriched KEGG pathways with ermineR. 

#### FGSEA KEGG pathways

```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysKEGG <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysKEGG)` up-regulated and `r nrow(downPathwaysKEGG)` down-regulated Reactome pathways with fgsea.

```{r}
upPathwaysKEGG %>% kable()
```
```{r}
upPathwaysKEGG109658_096 <- upPathwaysKEGG
save(upPathwaysKEGG109658_096, file="upPathwaysKEGG109658_096.Rdata")
```

### WikiPathways
#### ErmineR Wiki pathways

```{r}
enrichmentResultWiki <- precRecall(scores = ermineInputGeneScoresWiki,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways_changed_column.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

#enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultWiki$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  filter(CorrectedMFPvalue <= 0.05) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r}
#based on corrected MFPvalues 
sizeErmineWiki <- enrichmentResultWiki$results %>% filter(CorrectedMFPvalue <= 0.05)
```

There are `r nrow(sizeErmineWiki)` enriched Wiki pathways with ermineR

#### FGSEA Wiki pathways

```{r}
pathwaysWiki <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt")

#ES – enrichment score, same as in Broad GSEA implementation;
#NES – enrichment score normalized to mean enrichment of random samples of the same size;
fgseaRes <- fgsea(pathwaysWiki, scoresFGSEAWiki, minSize=15, maxSize=300, nperm=10000)

#up-regulated pathways
upPathwaysWiki <- fgseaRes %>% 
  arrange(padj) %>% filter(NES > 0) %>% filter(padj <= 0.05) %>% 
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))

#down-regulated pathways
downPathwaysWiki <- fgseaRes %>% 
  arrange(padj) %>% filter(NES < 0) %>% filter(padj <= 0.05) %>%
  dplyr::select(c("pathway", "padj", "ES", "NES", "nMoreExtreme"))
```

There are `r nrow(upPathwaysWiki)` up-regulated and `r nrow(downPathwaysWiki)` down-regulated Wiki pathways with fgsea

```{r}
upPathwaysWiki %>% kable()
```
```{r}
upPathwaysWiki109658_096 <- upPathwaysWiki
save(upPathwaysWiki109658_096, file="upPathwaysWiki109658_096.Rdata")
```
