---
title: "Pathway enrichment analysis"
author: "German Novakovskiy"
date: "August 17, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(knitr)
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(GEOquery)))
suppressWarnings(suppressMessages(library(data.table)))
suppressWarnings(suppressMessages(library(reshape2)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(cowplot)))
suppressWarnings(suppressMessages(library(limma)))
suppressWarnings(suppressMessages(library(tibble)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(edgeR)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(VennDiagram)))
suppressMessages(suppressWarnings(library(ermineR)))
suppressMessages(suppressWarnings(library(hgu133a.db)))
suppressMessages(suppressWarnings(library(ReactomePA)))
suppressMessages(suppressWarnings(library(clusterProfiler)))
suppressMessages(suppressWarnings(library(fgsea)))
```

Pathway enrichment analysis of individual definitive endoderm differentiation experiments (from GEO) and LINS data

## Analysis of pathways

KEGG pathways were dowloaded from here (http://www.broadinstitute.org/gsea/msigdb/collections.jsp#C2) in gmt format. Wikipathways were downloaded from the official website (http://data.wikipathways.org/20180810/gmt/), latest version August 10, 2018. Reactome pathways (gmt file) were dowloaded from here (https://reactome.org/download/current/ReactomePathways.gmt.zip).

### Analysis of REACTOME pathways

Reading data from file:
```{r}
x <- scan("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv", what = "", sep = "\n")
x <- strsplit(x, "[ \t]+")
max.col <- max(sapply(x, length))

## specify col.names as ?read.table suggests
cn <- paste("V", 1:max.col, sep = "")
reactome_pathways <- read.table("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv", fill = TRUE, 
                 col.names = cn, sep = '\t', quote = "")

reactome_pathways[1:5, 1:10] %>% kable()
```

There are `r nrow(reactome_pathways)` pathways in total.

Distribution of counts of genes in each pathway for REACTOME:
```{r}
calculateSize <- function(row_of_df){
  x <- row_of_df[3:length(row_of_df)]
  x <- x[!is.na(x)]
  return(sum(x != ""))
}

reactome_sizes_of_pathways <- apply(reactome_pathways, 1, calculateSize)

hist(reactome_sizes_of_pathways, breaks = 50, col = "red")
```

Sorting pathways by size:
```{r}
reactome_pathways_amount <- as.data.frame(reactome_pathways[,2])
colnames(reactome_pathways_amount) <- "Pathway"
reactome_pathways_amount$Size <- reactome_sizes_of_pathways

reactome_pathways_amount <- reactome_pathways_amount %>%
  arrange(desc(Size))
```

We see that are the largest pathways are very broad:
```{r}
head(reactome_pathways_amount, n = 10) %>% kable()
```

However, the smallest pathways contain just one gene:
```{r}
tail(reactome_pathways_amount, n = 10) %>% kable()
```

In our enrichment analysis we will consider only pathways that contain at least 15 genes and no more than 300 genes:
```{r}
reactome_selected_pathways <- reactome_pathways_amount %>% 
  filter(Size >= 15 & Size <= 300)
```

There are `r nrow(reactome_selected_pathways)` of such pathways.

### Analysis of KEGG pathways

Reading data from file:
```{r}
x <- scan("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt", what = "", sep = "\n")
x <- strsplit(x, "[ \t]+")
max.col <- max(sapply(x, length))

## specify col.names as ?read.table suggests
cn <- paste("V", 1:max.col, sep = "")
kegg_pathways <- read.table("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt", fill = TRUE, 
                 col.names = cn, sep = '\t', quote = "")

kegg_pathways <- kegg_pathways[,c(2,1,3:ncol(kegg_pathways))]
```
There are `r nrow(kegg_pathways)` pathways in total.

Distribution of counts of genes in each pathway for KEGG:
```{r}
kegg_sizes_of_pathways <- apply(kegg_pathways, 1, calculateSize)

hist(kegg_sizes_of_pathways, breaks = 50, col = "red")
```

Sorting pathways by size:
```{r}
kegg_pathways_amount <- as.data.frame(kegg_pathways[,2])
colnames(kegg_pathways_amount) <- "Pathway"
kegg_pathways_amount$Size <- kegg_sizes_of_pathways

kegg_pathways_amount <- kegg_pathways_amount %>%
  arrange(desc(Size))
```

The largest KEGG pathways:
```{r}
head(kegg_pathways_amount, n = 10) %>% kable()
```

Smallest pathways contain at least 10 genes:
```{r}
tail(kegg_pathways_amount, n = 10) %>% kable()
```

Again, we consider only pathways that contain at least 15 genes and no more than 300 genes:
```{r}
kegg_selected_pathways <- kegg_pathways_amount %>% 
  filter(Size >= 15 & Size <= 300)
```

There are `r nrow(kegg_selected_pathways)` of such pathways.

### Analysis of Wikipathways

Reading data from file:
```{r}
x <- scan("~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt", what = "", sep = "\n")
x <- strsplit(x, "[ \t]+")
max.col <- max(sapply(x, length))

## specify col.names as ?read.table suggests
cn <- paste("V", 1:max.col, sep = "")
wiki_pathways <- read.table("~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt", fill = TRUE, 
                 col.names = cn, sep = '\t', quote = "")

wiki_pathways <- wiki_pathways[,c(2,1,3:ncol(wiki_pathways))]
wiki_pathways[1:5, 1:10] %>% kable()
```

There are `r nrow(wiki_pathways)` pathways in total.

Distribution of counts of genes in each pathway for Wikipathways:
```{r}
wiki_sizes_of_pathways <- apply(wiki_pathways, 1, calculateSize)

hist(wiki_sizes_of_pathways, breaks = 50, col = "red")
```

Sorting pathways by size:
```{r}
wiki_pathways_amount <- as.data.frame(wiki_pathways[,2])
colnames(wiki_pathways_amount) <- "Pathway"
wiki_pathways_amount$Size <- wiki_sizes_of_pathways

wiki_pathways_amount <- wiki_pathways_amount %>%
  arrange(desc(Size))
```

Here are the largest pathways from Wiki:
```{r}
head(wiki_pathways_amount, n = 10) %>% kable()
```

Some pathways from Wiki contain just one gene.
```{r}
tail(wiki_pathways_amount, n = 10) %>% kable()
```

In our enrichment analysis we will consider only pathways that contain at least 15 genes and no more than 300 genes:
```{r}
wiki_selected_pathways <- wiki_pathways_amount %>% 
  filter(Size >= 15 & Size <= 300)
```

There are `r nrow(reactome_selected_pathways)` of such pathways.

## Analysis of GEO individual experiments

### Analysis of GSE109658

Load DE data:
```{r}
load("../analysis_of_public_data/GSE109658/DEgenes_0h_96h_109658.Rdata")
```

Sorted log Fold Changes give us a sorted list:
```{r}
ermineInputGeneScores <- DEgenes_0h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, absolute_logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(absolute_logFC)) %>% 
  column_to_rownames("gene")

save(ermineInputGeneScores, file = "ermineInputScores.Rdata")
head(ermineInputGeneScores, 10) %>% kable() # print the first few rows
```

#### Reactome pathways

```{r}
enrichmentResultReactome <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv",
                               minClassSize = 15,
                               maxClassSize = 300)

enrichmentResultReactome$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```
```{r}
enrichmentResultReactome$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```

#### Reactome pathways via ReactomPA

```{r}
#converting symbols to EntrezID
geneSymbols <- rownames(ermineInputGeneScores)
geneEntrez <- bitr(geneSymbols, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
```
```{r}
head(geneEntrez)
```
```{r}
ermineInputGeneScoresRPA <- ermineInputGeneScores %>%
  rownames_to_column("SYMBOL") %>%
  filter(SYMBOL %in% geneEntrez$SYMBOL)

inputGSEA <- merge(ermineInputGeneScoresRPA, geneEntrez, sort=FALSE)

inputGSEA <- inputGSEA %>%
  dplyr::select(absolute_logFC, ENTREZID) %>%
  column_to_rownames('ENTREZID')
```

Performing GSEA with the input:
```{r}
#test2 variable contains data with absolute log fold change
#and entrez id
test2 <- inputGSEA$absolute_logFC
names(test2) <- rownames(inputGSEA)

y <- gsePathway(test2, nPerm=10000,
                pvalueCutoff=0.05,
                pAdjustMethod="BH", verbose=FALSE, minGSSize = 15, maxGSSize = 300)

dotplot(y, showCategory=15)
```

#### KEGG pathways

```{r}
enrichmentResultKEGG <- precRecall(scores = ermineInputGeneScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

enrichmentResultKEGG$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultKEGG$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Pathway", "Corrected p-value", 
                                   "Corrected MF p-value"))
```

#### Wikipathways

```{r}
entrezScores <- as.data.frame(test2)
enrichmentResultWiki <- precRecall(scores = entrezScores,
                               scoreColumn = 1,
                               bigIsBetter = TRUE,
                               aspects = "B",
                               iterations = 10000,
                               geneSetDescription = NULL,
                               customGeneSets = "~/ESC_RNA_seq/pathway_enrichment_analysis/Wikipathways.gmt",
                               minClassSize = 15,
                               maxClassSize = 300)

enrichmentResultWiki$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

```{r}
enrichmentResultWiki$results %>% 
  dplyr::select(ID, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("ID", "Corrected p-value", 
                                   "Corrected MF p-value"))
```

#### Trying to run everything with fgsea

```{r}
# scores forfgsea
ermineInputGeneScoresFGSEA <- DEgenes_0h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  #mutate(absolute_logFC = abs(logFC)) %>% 
  dplyr::select(gene, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) %>% 
  column_to_rownames("gene")

scoresFGSEA <- ermineInputGeneScoresFGSEA$logFC
names(scoresFGSEA) <- rownames(ermineInputGeneScoresFGSEA)
```

Reactome (check the output columns!!!):
```{r}
pathwaysReactome <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv")

fgseaRes <- fgsea(pathwaysReactome, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#activated pathways
activPathwaysReactome <- fgseaRes %>% arrange(padj) %>% filter(NES > 0)
```

```{r}
#genes that are involved in TGF-beta pathway are enriched at the beginning of the list;
#means that this pathway is activated
plotEnrichment(pathwaysReactome[["R-HSA-9006936"]], scoresFGSEA) + labs(title="Signaling by TGF-beta family members")
```


KEGG (check the output columns!!!):
```{r}
pathwaysKEGG <- gmtPathways("~/ESC_RNA_seq/pathway_enrichment_analysis/KeggPathways.gmt")

fgseaRes <- fgsea(pathwaysKEGG, scoresFGSEA, minSize=15, maxSize=300, nperm=10000)

#activated pathways
activPathwaysKEGG <- fgseaRes %>% arrange(padj) %>% filter(NES > 0)
```

Results for KEGG also demonstrate enrichment of TGF-beta pathway:
```{r}
#genes that are involved in TGF-beta pathway are enriched at the beginning of the list;
#means that this pathway is activated
plotEnrichment(pathwaysKEGG[["KEGG_TGF_BETA_SIGNALING_PATHWAY"]], scoresFGSEA)+ labs(title="Signaling by TGF-beta family members")
```