---
title: "Pathway enrichment analysis"
author: "German Novakovskiy"
date: "August 17, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(knitr)
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(GEOquery)))
suppressWarnings(suppressMessages(library(data.table)))
suppressWarnings(suppressMessages(library(reshape2)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(cowplot)))
suppressWarnings(suppressMessages(library(limma)))
suppressWarnings(suppressMessages(library(tibble)))
suppressWarnings(suppressMessages(library(RColorBrewer)))
suppressWarnings(suppressMessages(library(edgeR)))
suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(VennDiagram)))
suppressMessages(suppressWarnings(library(ermineR)))
suppressMessages(suppressWarnings(library(hgu133a.db)))
suppressMessages(suppressWarnings(library(ReactomePA)))
suppressMessages(suppressWarnings(library(clusterProfiler)))
```

Pathway enrichment analysis of individual definitive endoderm differentiation experiments (from GEO) and LINS data

## Analysis of pathways

All pathways from KEGG/REACTOME for analysis with ErmineR should be dowloaded from here (http://www.broadinstitute.org/gsea/msigdb/collections.jsp#C2) in gmt format. 

### Analysis of REACTOME pathways

Reading data from file:
```{r}
x <- scan("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv", what = "", sep = "\n")
x <- strsplit(x, "[ \t]+")
max.col <- max(sapply(x, length))

## specify col.names as ?read.table suggests
cn <- paste("V", 1:max.col, sep = "")
reactome_pathways <- read.table("~/ESC_RNA_seq/pathway_enrichment_analysis/ReactomePathways.tsv", fill = TRUE, 
                 col.names = cn, sep = '\t', quote = "")

reactome_pathways[1:5, 1:10] %>% kable()
```

There are `r nrow(reactome_pathways)` pathways in total.

Distribution of counts of genes in each pathway for REACTOME:
```{r}
calculateSize <- function(row_of_df){
  x <- row_of_df[3:length(row_of_df)]
  x <- x[!is.na(x)]
  return(sum(x != ""))
}

sizes_of_pathways <- apply(reactome_pathways, 1, calculateSize)

hist(sizes_of_pathways, breaks = 50, col = "red")
```

Sorting pathways by size:
```{r}
pathways_amount <- as.data.frame(reactome_pathways[,2])
colnames(pathways_amount) <- "Pathway"
pathways_amount$Size <- sizes_of_pathways

pathways_amount <- pathways_amount %>%
  arrange(desc(Size))
```

We see that are the largest pathways are very broad:
```{r}
head(pathways_amount, n = 10) %>% kable()
```

However, the smallest pathways contain just one gene:
```{r}
tail(pathways_amount, n = 10) %>% kable()
```

In our enrichment analysis we will consider only pathways that contain at least 15 genes and no more than 300 genes:
```{r}
selected_pathways <- pathways_amount %>% 
  filter(Size >= 15 & Size <= 300)
```

There are `r nrow(selected_pathways)` of such pathways.

## Analysis of GEO individual experiments

### Analysis of GSE109658

Load DE data:
```{r}
load("../analysis_of_public_data/GSE109658/DEgenes_0h_96h_109658.Rdata")
```

Sorted log Fold Changes give us a sorted list:
```{r}
ermineInputGeneScores <- DEgenes_0h_96h_109658 %>% 
  rownames_to_column("gene") %>%
  dplyr::select(gene, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) %>% 
  column_to_rownames("gene")

save(ermineInputGeneScores, file = "ermineInputScores.Rdata")
head(ermineInputGeneScores, 10) %>% kable() # print the first few rows
```

```{r}
#enrichmentResult <- precRecall(scores = ermineInputGeneScores, 
#                               scoreColumn = 1, # column 1 is the scores 
#                               bigIsBetter = TRUE, # larger logFC should be ranked higher
#                               annotation = "Generic_human_custom", 
#                               aspects = "B", # look at only biological processes 
#                               iterations = 10000, # 10K sampling iterations so that results are stable
#                               geneSetDescription = "GO.xml",
#                               customGeneSets = "ReactomePathways.tsv",
#                               minClassSize = 15,
#                               maxClassSize = 300) 

#enrichmentResult$results %>% arrange(MFPvalue) %>% head(10) %>% kable()
```

While ErmineR is being fixed I will try GSEA for Reactome with specified bioconductor package: ReactomePA
```{r}
#converting symbols to EntrezID
geneSymbols <- rownames(ermineInputGeneScores)
geneEntrez <- bitr(geneSymbols, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

head(geneEntrez)
```
```{r}
ermineInputGeneScoresRPA <- ermineInputGeneScores %>%
  rownames_to_column("SYMBOL") %>%
  filter(SYMBOL %in% geneEntrez$SYMBOL)

inputGSEA <- merge(ermineInputGeneScoresRPA, geneEntrez, sort=FALSE)

inputGSEA <- inputGSEA %>%
  dplyr::select(logFC, ENTREZID) %>%
  column_to_rownames('ENTREZID')
```

Performing GSEA with the input:
```{r}
test2 <- inputGSEA$logFC
names(test2) <- rownames(inputGSEA)

y <- gsePathway(test2, nPerm=10000,
                pvalueCutoff=0.05,
                pAdjustMethod="BH", verbose=FALSE, minGSSize = 15, maxGSSize = 300)

dotplot(y, showCategory=15)
```

