---
title: "GSE52158 data analysis"
author: "German Novakovskiy"
date: "May 17, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data analysis of microarray data (Affymetrix Human Genome U133 Plus 2.0 Array, GPL570 platform) from that paper (PMID: 24412311)
```{r}
suppressMessages(suppressWarnings(library(limma)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(affy)))
suppressMessages(suppressWarnings(library(hgu133plus2.db)))
suppressMessages(suppressWarnings(library(tibble)))
suppressMessages(suppressWarnings(library(GEOquery)))
suppressMessages(suppressWarnings(library(reshape2)))
suppressMessages(suppressWarnings(library(ermineR)))
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(forcats)))
```

Loading the data:
```{r}
if (file.exists("GSE52158.Rdata")) {
  # if previously downloaded
  load("GSE52158.Rdata")
} else {
  # Get geo object that contains our data and phenotype information
  geo_obj <- getGEO("GSE52158", GSEMatrix = TRUE)
  geo_obj <- geo_obj[[1]]
  save(geo_obj, file = "GSE52158.Rdata")
}

show(geo_obj)
```
APS - anteriomost primitive streak
AFG - anterior foregut
PFG - posterior foregut
MHG - midgut/hindgut
```{r}
geo_metadata <- pData(geo_obj)[, c("organism_ch1", "title", colnames(pData(geo_obj))[grep("characteristics", 
                                                                                          colnames(pData(geo_obj)))])]

geo_metadata <- geo_metadata[,-3]
colnames(geo_metadata) <- c("organism", "sample", "description")
#only hESC, SR1 APS and SR1 DE 
geo_metadata <- geo_metadata[1:9,]
geo_metadata$cell_type <- as.factor(c(rep("ESC", 3), rep("APS", 3), rep("DE", 3)))
geo_metadata$cell_type <- relevel(geo_metadata$cell_type, "ESC")

geo_metadata
```

Expression data (data is RMA-normalized signal intensity (not log transformed) thus we have to log normalized it ):
```{r}
#data in RMA
data <- exprs(geo_obj)
data <- data[, rownames(geo_metadata)]
#filtering
keep.exprs <-rowSums(data > 20) >= 3 
data <- data[keep.exprs,]
log_data <- log(data+1)

#for plotting
hist(log_data)
```
```{r}
#for plotting
meltedExpressionMatrix <- log_data %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  melt(id = "gene") 

#for plotting
meltedExpressionMatrix %>% 
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
#all(colnames(data) == rownames(geo_metadata)) (TRUE)
designMatrix <- model.matrix(~cell_type, geo_metadata)

#linear fit 
cellTypeFit <- lmFit(log_data, designMatrix)
# run ebayes to calculate moderated t-statistics
#trend True - for accounting of low expressed probes
cellTypeFitEb <- eBayes(cellTypeFit)#, trend = T)

cellTypeRes <- decideTests(cellTypeFitEb, p.value = 0.05, lfc = 1)
summary(cellTypeRes)

```

For pathway analysis:

```{r}
#all(colnames(data) == rownames(geo_metadata)) (TRUE)
designMatrix2 <- model.matrix(~0 + cell_type, geo_metadata)

# construct the contrast matrix
contrastMatrix <- makeContrasts(
  APSvESC = cell_typeAPS - cell_typeESC,
  DEvESC = cell_typeDE - cell_typeESC,
  DEvAPS = cell_typeDE - cell_typeAPS,
  levels = designMatrix2
)

contrastMatrix %>% kable()

```
```{r}
#linear fit 
cellTypeFit2 <- lmFit(log_data, designMatrix2)
# fit the contrast using the original fitted model
contrastFit <- contrasts.fit(cellTypeFit2, contrastMatrix)

# apply eBayes() for moderated statistics
contrastFitEb <- eBayes(contrastFit)
```


```{r}
topProbesDE <- topTable(cellTypeFitEb, coef = "cell_typeDE", p.value = 0.05, lfc = 1, number = Inf) 
downTopProbesDE <- topProbesDE %>%
  rownames_to_column("probes") %>%
  filter(logFC < 0) #616

upTopProbesDE <- topProbesDE %>% 
  rownames_to_column("probes") %>%
  filter(logFC > 0) #775
```

Loading "golden" list of markers
```{r}
#get gene symbols
up_genes <- AnnotationDbi::select(hgu133plus2.db, upTopProbesDE$probes, c("SYMBOL"))
up_genes <- up_genes$SYMBOL
up_genes <- up_genes[!is.na(up_genes)]

down_genes <- AnnotationDbi::select(hgu133plus2.db, downTopProbesDE$probes, c("SYMBOL"))
down_genes <- down_genes$SYMBOL
down_genes <- down_genes[!is.na(down_genes)]
```

Which markers from golden list are also DE here?
```{r}
#check with golden list
up_golden_list <- scan("~/Papers/CMAP/up_genes_in_DE.grp", character(), quote = "")
up_golden_list[up_golden_list %in% up_genes]
```
Which are not?
```{r}
up_golden_list[!up_golden_list %in% up_genes]
```

The same for down regulated genes:
```{r}
#down_golden_list <- scan("~/Papers/CMAP/down_genes_in_DE.grp", character(), quote = "")
#down_golden_list[down_golden_list %in% down_genes]
```
```{r}
#down_golden_list[!down_golden_list %in% down_genes]
```
```{r}
#upRegulated52158 <- up_genes
#downRegulated52158 <- down_genes
#save(upRegulated52158, file="upRegulated52158.Rdata")
#save(downRegulated52158, file="downRegulated52158.Rdata")
```


# Gene set enrichment analysis of GSE75748

```{r}
topProbesDEAll <- topTable(cellTypeFitEb, coef = "cell_typeDE", number = Inf)

###for pathways
DEgenes_0h_96h <- topTable(contrastFitEb, number = Inf, coef = "DEvESC")
DEgenes_0h_24h <- topTable(contrastFitEb, number = Inf, coef = "APSvESC")
DEgenes_24h_96h <- topTable(contrastFitEb, number = Inf, coef = "DEvAPS")

#for pathway enrichment analysis
DEgenes_0h_96h_52158 <- DEgenes_0h_96h
DEgenes_0h_24h_52158 <- DEgenes_0h_24h
DEgenes_24h_96h_52158 <- DEgenes_24h_96h

save(DEgenes_0h_96h_52158, file = "DEgenes_0h_96h_52158.Rdata")
save(DEgenes_0h_24h_52158, file = "DEgenes_0h_24h_52158.Rdata")
save(DEgenes_24h_96h_52158, file = "DEgenes_24h_96h_52158.Rdata")
################################

x <- hgu133plus2SYMBOL
# Get the probe identifiers - gene symbol mappings
mapped_probes <- mappedkeys(x)
# Convert to a dataframe
genesym.probeid <- as.data.frame(x[mapped_probes])
head(genesym.probeid)
```
```{r}
probes_to_genes <- genesym.probeid %>%
  filter(probe_id %in% rownames(topProbesDEAll))

topProbes <- topProbesDEAll %>%
  rownames_to_column("probes") %>%
  filter(probes %in% probes_to_genes$probe_id)

probes_to_genes <- probes_to_genes %>% column_to_rownames('probe_id')
symbs <- probes_to_genes[topProbes$probes,]
topProbes$Symbol <- symbs

ermineInputProbeScores <- topProbes %>% 
  #as.data.frame() %>%
  #mutate(absolute_logFC = logFC) %>% 
  dplyr::select(probes, logFC) %>% 
  na.omit() %>% 
  as.data.frame() %>% 
  arrange(desc(logFC)) %>% 
  column_to_rownames("probes")

head(ermineInputProbeScores, 10)# %>% kable() # print the first few rows
```
```{r}
enrichmentResult <- precRecall(scores = ermineInputProbeScores, 
                               scoreColumn = 1, # column 1 is the scores 
                               bigIsBetter = TRUE, # larger logFC should be ranked higher
                               annotation = "GPL570", 
                               aspects = "B", # look at only biological processes 
                               iterations = 10000, # 10K sampling iterations so that results are stable
                               geneSetDescription = "../GSE109658/GO.xml") # use the GO XML file in current directory

enrichmentResult$results %>% arrange(MFPvalue) %>% head(10)
```
```{r}
enrichmentResult$results %>% 
  dplyr::select(Name, CorrectedPvalue, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  head(10) %>% 
  kable(align = "l", col.names = c("Biological Process", "Corrected p-value", 
                                   "Corrected MF p-value"))
```
```{r fig.width=15, fig.height=10}
Enrichment <- enrichmentResult$results
Enrichment$Name <- as.factor(Enrichment$Name)

Enrichment %>% 
  dplyr::select(Name, NumGenes, CorrectedMFPvalue) %>% 
  arrange(CorrectedMFPvalue) %>% 
  filter(CorrectedMFPvalue <= 5e-2) %>% 
  head(25) %>% 
  ggplot(aes(x = fct_reorder(Name, CorrectedMFPvalue), 
             y = NumGenes, fill = CorrectedMFPvalue)) +
  geom_col() +
  labs(title = "Biological Processes - GSE52158", 
       x = "", y = "Number of Genes") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip() +
  theme_light() 
```

# for CMAP (top 150 from up and 150 from down)

```{r}
load("../../common_upRegulated_52158_109658.Rdata")
load("../../common_downRegulated_52158_109658.Rdata")

genRef <- topProbes 

#taking upregulated genes
referenceUpregulated <- genRef %>% filter(Symbol %in% common_upRegulated_3)
referenceUpregulated <-  referenceUpregulated[with(referenceUpregulated, order(adj.P.Val)), ]
referenceUpregulatedLogSorted <- referenceUpregulated[with(referenceUpregulated, order(logFC, decreasing = T)), ]

#taking downregulated genes
referenceDownregulated <- genRef %>% filter(Symbol %in% common_downRegulated_3)
referenceDownregulated <-  referenceDownregulated[with(referenceDownregulated, order(adj.P.Val)), ]
referenceDownregulatedLogSorted <- referenceDownregulated[with(referenceDownregulated, order(logFC)), ]

#take 150 upregulated genes (by logFC)
first_up <- unique(referenceUpregulatedLogSorted$Symbol)[1:150]
fileConn<-file("~/Papers/CMAP/common_Upregulated_3_52158_LogSorted.grp")
writeLines(first_up, fileConn)
close(fileConn)

#take 150 downregulated genes (by logFC)
first_down <- unique(referenceDownregulatedLogSorted$Symbol)[1:150]
fileConn2<-file("~/Papers/CMAP/common_Downregulated_3_52158_LogSorted.grp")
writeLines(first_down, fileConn2)
close(fileConn2)
```
