---
title: "GSE52158 data analysis"
author: "German Novakovskiy"
date: "May 17, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data analysis of microarray data (Affymetrix Human Genome U133 Plus 2.0 Array) from that paper (PMID: 24412311)
```{r}
library(limma)
library(ggplot2)
library(dplyr)
library(affy)
library(hgu133plus2.db)
library(tibble)
library(GEOquery)
library(reshape2)
```

Loading the data:
```{r}
if (file.exists("GSE52158.Rdata")) {
  # if previously downloaded
  load("GSE52158.Rdata")
} else {
  # Get geo object that contains our data and phenotype information
  geo_obj <- getGEO("GSE52158", GSEMatrix = TRUE)
  geo_obj <- geo_obj[[1]]
  save(geo_obj, file = "GSE52158.Rdata")
}

show(geo_obj)
```
APS - anteriomost primitive streak
AFG - anterior foregut
PFG - posterior foregut
MHG - midgut/hindgut
```{r}
geo_metadata <- pData(geo_obj)[, c("organism_ch1", "title", colnames(pData(geo_obj))[grep("characteristics", 
                                                                                          colnames(pData(geo_obj)))])]

geo_metadata <- geo_metadata[,-3]
colnames(geo_metadata) <- c("organism", "sample", "description")
#only hESC, SR1 APS and SR1 DE 
geo_metadata <- geo_metadata[1:9,]
geo_metadata$cell_type <- as.factor(c(rep("ESC", 3), rep("APS", 3), rep("DE", 3)))
geo_metadata$cell_type <- relevel(geo_metadata$cell_type, "ESC")

geo_metadata
```

Expression data (data is RMA-normalized signal intensity (not log transformed) thus we have to log normalized it ):
```{r}
#data in RMA
data <- exprs(geo_obj)
data <- data[, rownames(geo_metadata)]
#filtering
keep.exprs <-rowSums(data > 20) >= 3 
data <- data[keep.exprs,]
log_data <- log(data+1)

#for plotting
hist(log_data)
```
```{r}
#for plotting
meltedExpressionMatrix <- log_data %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  melt(id = "gene") 

#for plotting
meltedExpressionMatrix %>% 
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
#all(colnames(data) == rownames(geo_metadata)) (TRUE)
designMatrix <- model.matrix(~cell_type, geo_metadata)

#linear fit 
cellTypeFit <- lmFit(log_data, designMatrix)
# run ebayes to calculate moderated t-statistics
#trend True - for accounting of low expressed probes
cellTypeFitEb <- eBayes(cellTypeFit)#, trend = T)

cellTypeRes <- decideTests(cellTypeFitEb, p.value = 0.05, lfc = 1)
summary(cellTypeRes)

```

```{r}
topProbesDE <- topTable(cellTypeFitEb, coef = "cell_typeDE", p.value = 0.05, lfc = 1, number = Inf) 
downTopProbesDE <- topProbesDE %>%
  rownames_to_column("probes") %>%
  filter(logFC < 0) #616

upTopProbesDE <- topProbesDE %>% 
  rownames_to_column("probes") %>%
  filter(logFC > 0) #775
```

Loading "golden" list of markers
```{r}
#get gene symbols
up_genes <- select(hgu133plus2.db, upTopProbesDE$probes, c("SYMBOL"))
up_genes <- up_genes$SYMBOL
up_genes <- up_genes[!is.na(up_genes)]

down_genes <- select(hgu133plus2.db, downTopProbesDE$probes, c("SYMBOL"))
down_genes <- down_genes$SYMBOL
down_genes <- down_genes[!is.na(down_genes)]
```

Which markers from golden list are also DE here?
```{r}
#check with golden list
up_golden_list <- scan("~/Papers/CMAP/up_genes_in_DE.grp", character(), quote = "")
up_golden_list[up_golden_list %in% up_genes]
```
Which are not?
```{r}
up_golden_list[!up_golden_list %in% up_genes]
```

The same for down regulated genes:
```{r}
down_golden_list <- scan("~/Papers/CMAP/down_genes_in_DE.grp", character(), quote = "")
down_golden_list[down_golden_list %in% down_genes]
```
```{r}
down_golden_list[!down_golden_list %in% down_genes]
```
```{r}
upRegulated52158 <- up_genes
downRegulated52158 <- down_genes
save(upRegulated52158, file="upRegulated52158.Rdata")
save(downRegulated52158, file="downRegulated52158.Rdata")
```

